Make any stupid plot you want

### 2021-02-19 draft/sections/math.tex
rework next + math on xi and hook

The datasets in figure~\ref{fig:base_example} have the same fiber of (temperature, time). In figure~\ref{fig:base_example_discrete} the fibers lie over discrete \dbase\ such that the records in the datasets in the fiber bundles are discrete. The same fiber in figure~\ref{fig:base_example_continuous} lies over a continuous interval \dbase such that the records are samples from a continuous function defined on \dbase. By encoding this continuity in the model as \dbase\, the data model now explicitly carries information about its structure such that the implicit assumptions of the visualization algorithms are now explicit. This in turn allows for building algorithms that can work with distributed or streaming data, since it has a common data access interface with a promise that the data exists.

@@

\subsubsection{Composition of Artists: +}
\begin{figure}
    \includegraphics[width=1\textwidth]{figures/math/addition_operator.png}
    \caption{$\dsection_1$ and $\dsection_2$ are distinct datasets passed through artists $\vartist_1$ and $\vartist_2$ to generate graphics $\gsection_1$ and $\gsection_2$. These graphics happen to be rendered to the same image, but otherwise have no intrinsinc link.}
    \label{fig:artist_plus}
\end{figure}
To build graphics that are the composites of multiple artists, we define a simple addition operator that is the disjoint union of fiber bundles \dtotal.  For example, in figure~\ref{fig:artist_plus} the scatter plot $\dtotal_1$ and the line plot $\dtotal_2$ have different \dbase that are mapped to separate \gbase. To fully display both graphics, the composite graphic $\vartist_1 + \vartist_2$ needs to include all records on both $\dbase_1$ and $\dbase_2$, which are the sections on the disjoint union $\dbase_1 \sqcup \dbase_2$. This in turn yields disjoint graphics $\gbase_1 \sqcup \gbase_2$ rendered to the same image. Constraints can be placed on the disjoint union such as that the fiber components need to have the same \vchannel position encodings or that the position \vsection need to be in a specified range. There are situations where $\dbase_2 \hookrightarrow \dbase_1$ that underpin more complex, especially interactive, visualizations; these cases require defining a more complex addition operator that is out of scope for this work.

>>>>

The datasets in figure~\ref{fig:base_example} have the same fiber of (temperature, time). In figure~\ref{fig:base_example_discrete} the fibers lie over discrete \dbase\ such that the records in the datasets in the fiber bundles are discrete. The same fiber in figure~\ref{fig:base_example_continuous} lies over a continuous interval \dbase\ such that the records are samples from a continuous function defined on \dbase. By encoding this continuity in the model as \dbase\, the data model now explicitly carries information about its structure such that the implicit assumptions of the visualization algorithms are now explicit. This in turn allows for building visualizations that can work with distributed or streaming data, since it has a common data access interface with a promise that the data exists. The \vindex\ mapping is also used by interactive visualization components to look up the data associated with a region on screen.  One example is to fill in details in a hover tooltip, another is to convert region selection (such as zooming) on \gbase\ to a query on the data to access the corresponding record components on \dbase.

@@

\subsubsection{Composition of Artists: +}
\begin{figure}
    \includegraphics[width=1\textwidth]{figures/math/addition_operator.png}
    \caption{$\dsection_1$ and $\dsection_2$ are distinct datasets passed through artists $\vartist_1$ and $\vartist_2$ to generate graphics $\gsection_1$ and $\gsection_2$. These graphics happen to be rendered to the same image, but otherwise have no intrinsinc link.}
    \label{fig:artist_plus}
\end{figure}
To build graphics that are the composites of multiple artists, we define a simple addition operator that is the disjoint union of fiber bundles \dtotal.  For example, in figure~\ref{fig:artist_plus} the scatter plot $\dtotal_1$ and the line plot $\dtotal_2$ have different \dbase that are mapped to separate \gbase. To fully display both graphics, the composite graphic $\vartist_1 + \vartist_2$ needs to include all records on both $\dbase_1$ and $\dbase_2$, which are the sections on the disjoint union $\dbase_1 \sqcup \dbase_2$. This in turn yields disjoint graphics $\gbase_1 \sqcup \gbase_2$ rendered to the same image. Constraints can be placed on the disjoint union such as that the fiber components need to have the same \vchannel position encodings or that the position \vsection need to be in a specified range. There is a second type of composition where $\dtotal_1$ and $\dtotal_2$ share a base space $\dbase_2 \hookrightarrow \dbase_1$ such that the the artists can be considered to be acting on different components of the same section. This type of composition is important for creating visualizations where elements need to update together in a consistent way, such as multiple views \cite{alboRadarComparativeEvaluation2016a, hullmanKeeping2018} and brush-linked views\cite{beckerBrushingScatterplots1987,bujaInteractiveData1991}.

### 2021-02-19 draft/sections/next.tex
rework next + math on xi and hook

\subsection{Future Work}
While the model and prototype demonstrate that generation of simple marks from the data, there is a lot of work left to develop a model that underpins a minimally viable library. Foremost is implementing a data object that encodes data with a 2D continous topology and an artist that can consume data with a 2D topology to visualize the image\cite{toryRethinkingVisualizationHighLevel2004,haber1990visualization,hansen2011visualization} and also encoding a separate heatmap\cite{wilkinsonHistoryClusterHeat2009,loua1873atlas} artist that consumes 1D discrete data.  A second important proof of concept artist is a boxplot\cite{wickham40YearsBoxplots2011} because it is a graphic that assumes computation on the data side and the glyph is built from semantically defined components and a list of outliers.

The model supports simple composition of glyphs by overlaying glyphs at the same position, but more work is needed to define an operator where the fiber bundles have shared $\gbase_2 \hookrightarrow \gbase_1$ such that fibers could be pulled back over the subset. This complex operator is necessary for building semantically meaningful components such as interactive visualizations that update on shared \gbase, such as brush-linked views\cite{beckerBrushingScatterplots1987,bujaInteractiveData1991} and verifying constraints in multiple view systems \cite{hullmanKeeping2018}. The models simple addition supports axes as standalone artists with overlapping visual position encoding, but the complex operator would allow for keeping labels consistent with text when for example the horizontal and vertical positions of the data are changed. In summary, the proposed scope of work for the dissertation is
\begin{itemize}
    \item expansion of the mathematical framework to include complex addition
    \item mathematical formulation of a graphic with axes labeling
    \item implementation of data with 2D continuous topology and a compatible artist
    \item provisional mathematics and implementation of user level composite artists
    \item proof of concept domain specific user facing library 
\end{itemize}
Additionally, implementing the complex addition operator would allow for developing a mathematical formalism and prototype of how interactivity would work in this model. Other potential tasks for future work is implementing a data object for a non-trivial fiber bundle and exploiting the models section level formalism to build distributed data source models and concurrent artists. This could be pushed further to integrate with topological\cite{heineSurveyTopologybasedMethods2016} and functional \cite{ramsayFunctionalDataAnalysis2006a} data analysis methods. While this paper formulates visualization in terms of monoidal action homomorphisms between fiberbundles, the model lends itself to a categorical formulation\cite{fongInvitationAppliedCategory2019,milewskiCategoryTheoryProgrammers} that could be further explored.

>>>>

\subsection{Future Work}
While the model and prototype demonstrate that generation of simple marks from the data, there is a lot of work left to develop a model that underpins a minimally viable library. Foremost is implementing a data object that encodes data with a 2D continous topology and an artist that can consume data with a 2D topology to visualize the image\cite{toryRethinkingVisualizationHighLevel2004,haber1990visualization,hansen2011visualization} and also encoding a separate heatmap\cite{wilkinsonHistoryClusterHeat2009,loua1873atlas} artist that consumes 1D discrete data.  A second important proof of concept artist is a boxplot\cite{wickham40YearsBoxplots2011} because it is a graphic that assumes computation on the data side and the glyph is built from semantically defined components and a list of outliers. The model supports simple composition of glyphs by overlaying glyphs at the same position, but more work is needed to define an operator where the fiber bundles have shared $\gbase_2 \hookrightarrow \gbase_1$ such that fibers could be pulled back over the subset. While the model's simple addition supports axes as standalone artists with overlapping visual position encoding, the complex operator would allow for binding together data that needs to be updated together.
In summary, the proposed scope of work for the dissertation is
\begin{itemize}
    \item expansion of the mathematical framework to include complex addition
    \item implementation of artist with explicit \vindex\
    \item specification of interactive visualization
    \item mathematical formulation of a graphic with axes labeling
    \item provisional mathematics and implementation of user level composite artists
    \item proof of concept domain specific user facing library 
\end{itemize}
Additionally, implementing the complex addition operator and explicit graphic to data maps would allow for developing a mathematical formalism and prototype of how interactivity would work in this model. Other potential tasks for future work is implementing a data object for a non-trivial fiber bundle and exploiting the models section level formalism to build distributed data source models and concurrent artists. This could be pushed further to integrate with topological\cite{heineSurveyTopologybasedMethods2016} and functional \cite{ramsayFunctionalDataAnalysis2006a} data analysis methods. While this paper formulates visualization in terms of monoidal action homomorphisms between fiberbundles, the model lends itself to a categorical formulation\cite{fongInvitationAppliedCategory2019,milewskiCategoryTheoryProgrammers} that could be further explored.
